---
output:
  html_document: default
---
---
title: "HWK 1"
author: "Dingxian Cao"
date: "September 9, 2015"
output: html_document
---  

# 1  

## 1.a
```{r}
suppressMessages(library(quantmod,quietly = T))
suppressMessages(getSymbols("^GSPC",from = "2012/01/03", to= "2015/09/08"))
head(GSPC)
day_return<-dailyReturn(Ad(GSPC),type = "arithmetic")#use adjusted
chartSeries(day_return,type="line",name="simple daily return of S&P500",theme = chartTheme("white"))
``` 

## 1.b
```{r}
suppressMessages(library(fBasics,quietly = T))
brief <-function(x) return(c(mean(x),sd(x),skewness(x),kurtosis(x)-3,min(x),max(x)))
return_summary<-brief(day_return)
names(return_summary)<-c("mean","sd","skewness","excess kurtosis","min","max")
data.frame(return_summary) 
```

## 1.c  
```{r}
plot(density(day_return),main="empirical density")
curve(dnorm(x,mean = mean(day_return),
            sd = sd(day_return)),
      xlim=c(-0.04,+0.04),
      lty=2,add=T)
legend("topright",lty=c(1,2),c("empirical","theory"))
normalTest(day_return,method="jb") 
```  

According to the empirical density curve compared to the normal density curve and the JB-test, we can say that the simple return of S&P doesn't have a normal distribution.

## 1.d
```{r}
log_return<-log(day_return+1)
head(log_return)
return_summary<-brief(log_return)
names(return_summary)<-c("mean","sd","skewness","excess kurtosis","min","max")
data.frame(return_summary)  
```

## 1.e 

```{r}
t.test(log_return) 
```
From the test, we can say that the mean of the log return of S&P is zero.

## 1.f
```{r}
abs(skewness(log_return)/sqrt(6/length(log_return)))>qnorm(0.05/2)
```
I compute the  Skewness(N(0,6/T)) statistic and it turns out that under the 5% significance level, the log return of S&P is not symmetric.

## 1.g 

```{r}
plot(density(log_return),main="empirical density",col="red")
curve(dnorm(x,mean = mean(log_return),
            sd = sd(log_return)),
      xlim=c(-0.04,+0.04),
      lty=2,col="blue",add=T)
legend("topright",lty=c(1,2),col=c("red","blue"),c("empirical","theory"))

(kurtosis(log_return)-3)/sqrt(24/length(log_return))>qnorm(0.05/2)
```

According to the empirical density curve compared to the normal density curve and the Kurtosis(N(0,24/T)) statistic, we can say that the **tail behavior** of the log return of S&P is normal!

# 2 

## 2.a
```{r} 
suppressMessages(getFX("USD/CNY", from="2012-01-03", to = "2015-09-08"))
head(USDCNY)
log_return<-dailyReturn(USDCNY,type = "log")
chartSeries(log_return,type="line",name="log return of USD/CNY",theme = chartTheme("white"))
```

## 2.b
```{r}
return_summary<-brief(log_return)
names(return_summary)<-c("mean","sd","skewness","excess kurtosis","min","max")
data.frame(return_summary) 
```

## 2.c
```{r}
plot(density(log_return),main="empirical density")
curve(dnorm(x,mean = mean(log_return),
            sd = sd(log_return)),
      xlim=c(-0.04,+0.04),
      lty=2,add=T)
legend("topright",lty=c(1,2),c("empirical","theory"))
```

## 2.d
```{r}
t.test(log_return)
```
the t test turns out that the mean of the daily log return of the exchange rate is 0.

## 2.e  
```{r}
normalTest(log_return,method="jb") 
```
According to the JB test, the daily log return of the exchange is not normally distributed.
