---
title: "hwk_601_1"
author: "Dingxian Cao"
date: "September 30, 2015"
output: word_document
---

# ex2
take k=9074463432 mod 35

```{r}
options(digits = 4)
setwd("/Users/CDX/WISC_R_HWK/Regression")
suppressMessages(library(broom))
library(car)
ext2<-read.table("HWK2_Ext2.txt")
(k<-9074463432 %% 35)
d<-ext2[((k-1)*90+1):(k*90),]
dim(d)

d1<-d[1:30,]
d2<-d[31:60,]
d3<-d[61:90,]
```

## data 1
```{r}
knitr::kable(tidy(cor(d1)))
scatterplotMatrix(d1,spread=F,main="data1")
```

>From the scatter plot and the covariance matrix we can see that there are no obvious relationship between V1 and V3, so there shouldn't be a response variable between the two.    

>But they both have a clear relationship with V2.So,I believe there is V2 is the reponse variable.   

```{r}
d11<-as.data.frame(d1)
names(d11)<-c("x1","y","x2")
```


#### data1 for model 1
```{r}
form1<- as.formula("y ~ 1+x1+x1:x2")
m1<-lm(form1,data=d11);knitr::kable(tidy(m1))
```

#### data1 for model2 
```{r}
form2<-as.formula("1/y ~ 1+x1+x2")
m2<-lm(form2,data=d11);knitr::kable(tidy(m2))
```

#### data1 for model3  

```{r}
form3<-as.formula("y ~ 0+x1+x2+I(x1^2)")
m3<-lm(form3,data=d11);knitr::kable(tidy(m3))
```  

>According to the P value of each coefficients, the 2nd model is obviously the best fitted.  

## data 2
```{r}
knitr::kable(tidy(cor(d2)))
scatterplotMatrix(d2,spread=F,main="data2")
```


>From the scatter plot and the covariance matrix we can see that there are no obvious relationship between V1 and V3, so there shouldn't be a response variable between the two.    

>But they both have a clear relationship with V2.So,I believe there is V2 is the reponse variable.   

```{r}
d21<-as.data.frame(d2)
names(d21)<-c("x2","y","x1")
```


#### data2 for model 1
```{r}
form1<- as.formula("y ~ 1+x1+x1:x2")
m1<-lm(form1,data=d21);knitr::kable(tidy(m1))
```

#### data2 for model3  

```{r}
form3<-as.formula("y ~ 0+x1+x2+I(x1^2)")
m3<-lm(form3,data=d21);knitr::kable(tidy(m3))
```  

>According to the P value of each coefficients, the 1st obviously the best fitted.  

## data 3  

```{r}
knitr::kable(tidy(cor(d3)))
scatterplotMatrix(d3,spread=F,main="data3")
```  

>From the scatter plot and the covariance matrix we can see that there are no obvious relationship between V1 and V3, so there shouldn't be a response variable between the two.    

>But they both have a clear relationship with V2.So,I believe there is V2 is the reponse variable.   

```{r}
d31<-as.data.frame(d3)
names(d31)<-c("x1","y","x2")
```

#### data3 for model3  

```{r}
form3<-as.formula("y ~ 0+x1+x2+I(x1^2)")
m3<-lm(form3,data=d31);knitr::kable(tidy(m3))
```  

>According to the P value of each coefficients, the 3rd model is obviously the best fitted.  

# ex3   

```{r}
#install.packages("SenSrivastava")
library(SenSrivastava)
suppressMessages(library(dplyr))
y_range<-range(E1.19$Price)
x_range<-range(E1.19$P)
group<-split(E1.19,E1.19$B)
par(mfrow=c(1,1))
plot(group$c[c(2,1)],xlim=x_range,ylim=y_range,pch=10,main="type = c")
plot(group$p[c(2,1)],xlim=x_range,ylim=y_range,pch=11,main="type = p")

library(ggplot2)
ggplot(E1.19,aes(x=P,y = Price,group=B,colour=B,shape=B))+ xlab("#page")+geom_line()
```  

>Since their axes are the same, I don't think there is any difference betweent individual and overlaid graph helping me explore the data.

>I personally don't want to distroy the structure of original data. But since the x variable varies wider than y variable, I may suggest to scale the x variable by letting each number be divided by their maximum.

```{r}
qplot(P,Price,data=E1.19)+
  geom_smooth(aes(colour=B),method="lm",se=F)+
  facet_wrap( ~ B,scale="free_y")+
  labs(title="Key Graph")

```

This picture clearly shows the linear relationship between price and page's volume in terms of two different kind of books.

>review: I don't think there is meaning in identifying the x1 and x2 by fitting the models.

# ex4  

```{r}
grocery<-read.table("grocery_retailer.txt",header = T)
plot(grocery)
```  

>I can get the infomation of pair-wise correlation which can help we varify the signs of coefficients.

```{r}
m<-lm(Y~X1+X2+X3,data = grocery)
knitr::kable(tidy(m))
```  

>The coefficient of X3 means that if there happens to be a holiday the total labor hours will be about 623 hours more than usual.

```{r}
plot(resid(m),type="l",main="residuals")
abline(h=mean(resid(m)))

layout(matrix(c(1,2,3,4),2,2,byrow = T))
plot(x=grocery$X1,y=resid(m),main="X1 vs Residuals")
plot(x=grocery$X2,y=resid(m),main="X2 vs Residuals")
plot(x=grocery$X3,y=resid(m),main="X3 vs Residuals")
```

>Apparently,there is a pattern between the residuals and independent variable X3.The variance of the error item can change along with X3. So the Gauss-Markov assuption is violated.

