---
title: "hwk_601_1"
author: "Dingxian Cao"
date: "September 30, 2015"
output: html_document
---

# ex2
take k=2 

```{r}
setwd("/Users/CDX/WISC_REG_WK/")
ext2<-read.table("HWK2_Ext2.txt")
k<-2 
d<-ext2[((k-1)*90+1):(k*90),]
dim(d)

d1<-d[1:30,]
d2<-d[31:60,]
d3<-d[61:90,]

plot(d)
```
From the scatter plot we can see that there are no obvious relationship between V1 and V3, so there shouldn't be a response variable between the two.    

But they both have a clear relationship with V2.So,I believe there is V2 is the reponse variable.   

```{r}
plot(cbind(d,d$V1^2,d$V3^2,d$V1*d$V3)[-c(1,3)])
```

Since we don't know the true model, we cannot use the model to identify the explantory variables. So I first assume X1=V1.

```{r}
suppressMessages(library(broom))
options(digits = 4)
d11<-as.data.frame(d1)
names(d11)<-c("x1","y","x2")
form1<- as.formula("y ~ 1+x1+x1:x2")
m1<-lm(form1,data=d11);knitr::kable(tidy(m1))

d21<-as.data.frame(d2)
names(d21)<-c("x1","y","x2")
form2<-as.formula("1/y ~ 1+x1+x2")
m2<-lm(form2,data=d21);knitr::kable(tidy(m2))

d31<-as.data.frame(d3)
names(d31)<-c("x1","y","x2")
form3<-as.formula("y ~ 0+x1+x2+I(x1^2)")
m3<-lm(form3,data=d31);knitr::kable(tidy(m3))

```  

>According to the P value of each coefficients, the third model is obviously the best fitted.  

# ex3   

```{r}
#install.packages("SenSrivastava")
library(SenSrivastava)
suppressMessages(library(dplyr))
y_range<-range(E1.19$Price)
x_range<-range(E1.19$P)
group<-split(E1.19,E1.19$B)
par(mfrow=c(2,1))
plot(group$c[c(2,1)],xlim=x_range,ylim=y_range,pch=10,main="type = c")
plot(group$p[c(2,1)],xlim=x_range,ylim=y_range,pch=11,main="type = p")

library(ggplot2)
ggplot(E1.19,aes(x=P,y = Price,group=B,colour=B,shape=B))+ xlab("#page")+geom_line()
```
Since their axes are the same, I don't think there is any difference betweent individual and overlaid graph helping me explore the data.

I personally don't want to distroy the structure of original data. But since the x variable varies wider than y variable, I may suggest to scale the x variable by letting each number be divided by their maximum.

```{r}
qplot(P,Price,data=E1.19)+
  geom_smooth(aes(colour=B),method="lm",se=F)+
  facet_wrap( ~ B,scale="free_y")+
  labs(title="Key Graph")

```

This picture clearly shows the linear relationship between price and page's volume in terms of two different kind of books.


# ex4  

```{r}
grocery<-read.table("grocery_retailer.txt",header = T)
plot(grocery)
```
I can get the infomation of pair-wise correlation which can help we varify the signs of coefficients.

```{r}
m<-lm(Y~X1+X2+X3,data = grocery)
knitr::kable(tidy(m))
```  

The coefficient of X3 means that if there happens to be a holiday the total labor hours will be about 6 hours more than usual.

```{r}
plot(resid(m),type="l",main="residuals")
abline(h=mean(resid(m)))

layout(matrix(c(1,2,3,4),2,2,byrow = T))
plot(x=grocery$X1,y=resid(m),main="X1 vs Residuals")
plot(x=grocery$X2,y=resid(m),main="X2 vs Residuals")
plot(x=grocery$X3,y=resid(m),main="X3 vs Residuals")
```

Apparently,there is a pattern between the residuals and independent variable X3. So the Gauss-Markov assuption is violated.

