---
title: "327_work1"
author: "Dingxian Cao"
date: "November 10, 2015"
output: html_document
---
# STAT 327 Group work on Optimization

Please write the names and email addresses of your group members.
Here's an example:

* Name / email (@wisc.edu only): John Gillett / jgillett@wisc.edu

* Dingxian Cao / email (dcao28@wisc.edu)
* Mingyi Yang / email (myang263@wisc.edu): ...
* Rouyi Ding/ email (rding26@wisc.edu): ...

# Part 1: One-dimensional optimization

Graph the object's altitude vs. time.

```{r}
z0=100;m=80;c=15;v0=55;g=9.81

z <- function(t){
  z0+(m/c)*(v0+m*g/c)*(1- exp(-c*t/m)) - m*g*t/c
}

curve(z,from = 0, to = 13)
abline(h=0)
```

Find the time at which the object strikes the ground.

```{r}
f<-function(x){
  abs(z(x))
}
optimize(f,c(0,20))
```

Find the object's maximum height.

```{r}
f<-function(x){
  -(z(x))
}
opt<-optimize(f,interval = c(0,14))
opt$minimum

```  

Find the time at which the object reaches its maximum height.

```{r}
z(opt$minimum)
#abline(v=opt$minimum)
```

# Part 2: Multi-dimensional optimization

## function plot
```{r}
f = function(x,y) { # Define function z = f(x, y).
  return(7.9+0.13*x+0.21*y-0.05*(x^2)-0.016*(y^2)-0.007*x*y)
}

# Plot function, z = f(x, y) over the region -pi/2 <= x, y <= pi/2.
grid.x = seq(-10, 10, length.out=20)
grid.y = seq(0,20,length.out=20)
grid.z = outer(grid.x, grid.y, f)
persp.out <- persp(grid.x, grid.y, grid.z, theta = 45, phi=60,
  main="Gradient descent", ticktype="detailed",
  xlab="x", ylab="y", zlab="z")
```

## Use `gradient.descent()` to find the peak.
```{r}
gradient.descent = function(par, gr1, gr2, gamma=.1, epsilon=.02, n=100, verbose=FALSE, ...) {
  for (i in seq_len(n)) {
    gradient1 = gr1(par[1],par[2])
    par[1] = par[1] - gamma * gradient1
    gradient2 = gr2(par[1],par[2])
    par[2] = par[2] - gamma * gradient2
    gradient.size = sum(abs(min(gradient1,gradient2)))
    
    if (verbose) {
      cat(sep="", "i=", i, ", par=c(", paste(signif(par, 4), collapse=","),
          "), gradient=c(", paste(signif(gradient1, 4),signif(gradient2, 4), collapse=","),
          "), size=", signif(gradient.size, 4), "\n")
    }
    if (gradient.size < epsilon) {
      break
    }
  }
  return(par)
}

gradient.x = function(x, y) {0.13-0.1*x-0.007*y}
gradient.y = function(x, y) {0.21-0.032*y-0.007*x}
(p<-gradient.descent(par=c(1,7), gr1=gradient.x, gr2=gradient.y, gamma=.01, epsilon=.01, n=100,verbose=FALSE))
z<-f(p[1],p[2])
persp.out <- persp(grid.x, grid.y, grid.z, theta = 45, phi=60,
  main="Gradient descent", ticktype="detailed",
  xlab="x", ylab="y", zlab="z")
points(trans3d(x=p[1], y=p[2], z=z, pmat=persp.out), col="red", pch = 19)
```

## Use `optim()` with `method=Nelder-Mead` to find the peak.
```{r}
grr <- function(x) { ## Gradient of 'fr'
    x1 <- x[1]
    x2 <- x[2]
    grad <- c(0.13-0.1*x1-0.007*x2,
       0.21-0.032*x2-0.007*x1)
    grad <- -grad
    return(grad)
}

f1 = function(x) { # Define function z = f(x, y).
  x1 <- x[1]
  x2 <- x[2]
  c<- -(7.9+0.13*x1+0.21*x2-0.05*(x1^2)-0.016*(x2^2)-0.007*x1*x2)
  return(c)
}

optim(par = c(1,7),fn = f1, grr, method ="Nelder-Mead" )
system.time(optim(par = c(1,7),fn = f1, grr, method ="Nelder-Mead" ))
```

## Use `optim()` with `method=BFGS` to find the peak.
```{r}
optim(par = c(1,7),fn = f1, grr,method ="BFGS" )
system.time(optim(par = c(1,7),fn = f1, grr,method ="BFGS" ))
```

How many calls did `optim()` make in each case? Which method would you
expect to be faster?

>Since the BFGS method only makes 18 calls for the function and gradient while Nelder-Mead method makes 43 calls for the function, so we think the method BFGS is better. 




